There are different types of abstraction techniques that can be applied to hybrid automata, we will discuss these in the sections below. What these techniques have in common is that they construct a discrete \emph{transition system} and they show that the abstraction preserves the properties of the original system. Some of the techniques use the properties to build partitions of the hybrid system, which then are abstracted.

\begin{defi}[Transition System \cite{Alur2000}]
A \emph{transition system} $\mathcal{T}=(Q,Q_{0},\Pi,t,\models)$ consists of
\begin{description}
    \item[$Q$]{the finite set of states;}
    \item[$Q_{0}\subseteq Q$]{the set of initial states;}
    \item[$\Pi$]{the set of propositions;}
    \item[$t \subseteq Q\times Q$]{a transition function;}
    \item[$\models\subseteq Q\times\Pi$]{a satisfaction function.}
\end{description}
\end{defi}

The \emph{properties} of a system are desired specifications, such as avoiding an unsafe set of states, that the system satisfies. Properties are used to formally verify the system and are described using logical formulae built from atomic propositions.

We denote by $\llbracket \pi \rrbracket = \{ q\in Q : q\models \pi\}$ the set of states satisfying the property $\pi$.

\begin{defi}[Linear Temporal Logic]
The formulae of \emph{linear temporal logic} (LTL) are defined recursively
\begin{description}
    \item[Propositions]{Every atomic proposition $\pi$ is a formula.}
    \item[Formulas]{If $\phi_{1}$ and $\phi_{2}$ are formulae then so are the following;
        \[
        \phi_{1} \lor \phi_{2},\ \ \phi_{1} \land \phi_{2}, \ \ \lnot \phi_{1}, \ \
        \phi_{1} \Rightarrow \phi_{2},\ \ \phi_{1} \Leftrightarrow \phi_{2}, \ \ \bigcirc \phi_{1}, \ \ \phi_{1} \until \phi_{2}, \ \ \lozenge \phi_{1}, \ \ \square \phi_{1}.
        \]}
\end{description}
\end{defi}
Where the formulae are interpreted over infinite words of propositions. Let $w=\Pi_{0}\Pi_{1}\ldots$ be a word with $\Pi_{i}$ being a set of propositions, then the satisfaction of a proposition $\pi$ at position $i\in \mathbb{N}$ of $w$ is denoted by $(w,i)\models_{L} \pi$ and holds if and only if $\pi \in \Pi_{i}$. Note that $\models_{L}$ is not the satisfaction relation $\models$ of a transition system.

\begin{defi}[Semantics of LTL]
Let $\phi_{1},\phi_{2}$ be LTL formulae, $w$ a word of propositions, and $i\in\mathbb{N}$. The semantics of any LTL formula is defined recursively as
\begin{itemize}
    \item{$(w,i)\models_{L}\phi_{1}\lor\phi_{2}$ if either $(w,i)\models_{L}\phi_{1}$ or $(w,i)\models_{L}\phi_{2}$;}
    \item{$(w,i)\models_{L}\phi_{1}\land\phi_{2}$ if both $(w,i)\models_{L}\phi_{1}$ and $(w,i)\models_{L}\phi_{2}$;}
    \item{$(w,i)\models_{L}\lnot\phi_{1}$ if $(w,i)\not\models_{L}\phi_{1}$;}
    \item{$(w,i)\models_{L}\phi_{1}\Rightarrow\phi_{2}$ if $(w,i)\models_{L}\lnot\phi_{1} \land \phi_{2}$;}
    \item{$(w,i)\models_{L}\phi_{1}\Leftrightarrow\phi_{2}$ if $(w,i)\models_{L}(\phi_{1}\Rightarrow\phi_{2})\land(\phi_{2}\Rightarrow\phi_{1})$;}
    \item{$(w,i)\models_{L}\bigcirc\phi_{1}$ if $(w,i+1)\models_{L}\phi_{1}$;}
    \item{$(w,i)\models_{L}\phi_{1}\until\phi_{2}$ if there is $j\geq i$ such that $(w,j)\models_{L}\phi_{2}$ and for all $i\leq k < j$ $(w,k)\models_{L}\phi_{1}$;}
    \item{$(w,i)\models_{L}\lozenge\phi_{1}$ if $(w,i)\models_{L}true\until\phi_{1}$;}
    \item{$(w,i)\models_{L}\square\phi_{1}$ if $(w,i)\models_{L}\lnot\lozenge\lnot\phi_{1}$.}
\end{itemize}
\end{defi}

The \emph{next} operator $\bigcirc\phi$ holds for a word $\Pi_{0}\Pi_{1}\Pi_{2}\ldots$ if and only if $\phi$ holds for $\Pi_{1}\Pi_{2}\ldots$. The \emph{until} operator $\phi_{1}\until\phi_{2}$ describes that $\phi_{1}$ is true until $\phi_{2}$ becomes true. The \emph{eventually} operator $\lozenge\phi$ indicates that $\phi$ will be true at some point in $w$. The \emph{always} operator $\square\phi$ describes that $\phi$ must always be true in a word.

We can extend LTL to include existential quantifiers. These types of quantifiers are used in property formulations of transition systems and the satisfaction relation of correlates to the satisfaction relation of this extended logic.

\begin{defi}[Computational Tree Logic]
    The formulae of \emph{Computational temporal logic} (CTL) are defined recursively
\begin{description}
    \item[Propositions]{Every atomic proposition $\pi$ is a formula.}
    \item[Formulas]{If $\phi_{1}$ and $\phi_{2}$ are formulae then so are the following;
        \[
        \phi_{1} \lor \phi_{2},\ \ \phi_{1} \land \phi_{2}, \ \ \lnot \phi_{1}, \ \
        \phi_{1} \Rightarrow \phi_{2},\ \ \phi_{1} \Leftrightarrow \phi_{2}, \ \
        \exists\bigcirc \phi_{1}, \ \ \phi_{1} \exists\until \phi_{2}, \ \
        \exists\lozenge \phi_{1}, \ \ \exists\square \phi_{1}, \ \ \forall\bigcirc \phi_{1}, \ \
        \forall\lozenge \phi_{1}, \ \ \forall\square \phi_{1}.
        \]}
\end{description}
\end{defi}

The difference between LTL and CTL semantics is that LTL is interpreted over words, whereas CTL is interpreted over trees generated by trajectories from a given state of a transition system. So a transition system $\mathcal{T}$ satisfies a proposition $\pi$ if there is a state $q_{0}\in Q$ in $\mathcal{T}$ such that $q_{0}\models \pi$.
\begin{defi}[Semantics of CTL]
Let $q_{0}\in Q$ be a state of a transition system $\mathcal{T}$, and $\phi_{1},\phi_{2}$ CTL formulae, then the semantics of CTL is defined as
\begin{itemize}
    \item{$q_{0}\models\phi_{1}\lor\phi_{2}$ if either $q_{0}\models\phi_{1}$ or $q_{0}\models\phi_{2}$;}
    \item{$q_{0}\models\phi_{1}\land\phi_{2}$ if both $q_{0}\models\phi_{1}$ and $q_{0}\models\phi_{2}$;}
    \item{$q_{0}\models\lnot\phi_{1}$ if $q_{0}\not\models\phi_{1}$;}
    \item{$q_{0}\models\phi_{1}\Rightarrow\phi_{2}$ if $q_{0}\models\lnot\phi_{1} \land \phi_{2}$;}
    \item{$q_{0}\models\phi_{1}\Leftrightarrow\phi_{2}$ if $q_{0}\models(\phi_{1}\Rightarrow\phi_{2})\land(\phi_{2}\Rightarrow\phi_{1})$;}
    \item{$q_{0}\models\exists\bigcirc\phi_{1}$ if $\exists q_{1}\in Q$ such that
    $t(q_{0})=q_{1}$ and $q_{1}\models \phi_{1}$;}
    \item{$q_{0}\models\phi_{1}\exists\until\phi_{2}$ if $\exists q_{0}q_{1}q_{2}\ldots$ a trajectory generated from $q_{0}$ such that $q_{i}\models \phi_{2}$ for some $i\geq 0$ and for all $0\leq j < i$, $q_{j}\models \phi_{1}$;}
    \item{$q_{0}\models\exists\lozenge\phi_{1}$ if $q_{0}\models true\ \exists\until \phi_{1}$;}
    \item{$q_{0}\models\exists\square\phi_{1}$ if $\exists q_{0}q_{1}q_{2}\ldots$ generated from $q_{0}$ such that $\forall i\geq 0$ $q_{i}\models\phi_{1}$;}

    \item{$q_{0}\models\forall\bigcirc\phi_{1}$ if $q_{0}\models\lnot\exists\bigcirc\lnot\phi_{1}$;}
    \item{$q_{0}\models\forall\lozenge\phi_{1}$ if $q_{0}\models\lnot\exists\square\lnot\phi_{1}$;}
    \item{$q_{0}\models\forall\square\phi_{1}$ if $q_{0}\models\lnot\exists\lozenge\lnot\phi_{1}$.}
\end{itemize}
\end{defi}

The temporal operators $\exists\bigcirc, \exists\until, \exists\lozenge$ and $\exists\square$ are called \emph{possibly next}, \emph{possibly always}, \emph{possibly until}, and \emph{possibly eventually} respectively and they refer to the existence of a trajectory starting at a given state. On the other hand, the operators $\forall\bigcirc$ (inevitably next), $\forall\lozenge$ (inevitably eventually) and $\forall\square$ (inevitably always) refer to all trajectories from a given state.

In general we can say that a transition system $\mathcal{T}$ \emph{satisfies} a CTL formula $\phi$ if some initial state of $\mathcal{T}$ satisfies $\phi$.

\subsection{General Verification Problems}
There are general questions that we can ask of a system which we are looking to be verified for example, we reach all possible states of the system.

\begin{prob}[Reachability Problem \cite{Henzinger2000}]
\label{prob:reach}
Given a hybrid automaton $\mathcal{A}$ and a proposition $\phi$. Is there a trajectory in the hybrid automaton that starts in an initial state and reaches a state $(l_{i},v)\in L\times\val(V)$ such that $(l_{i},v)\models\phi$?
\end{prob}

\begin{prob}[Emptiness Problem \cite{Henzinger2000}]
\label{prob:empty}
Given a hybrid automaton $\mathcal{A}$, is there a divergent trajectory starting in an initial state?
\end{prob}

%A \emph{timed trace} in a hybrid automaton is
%\begin{prob}[Timed Trace Inclusion Problem \cite{Henzinger2000}]
%\label{prob:time}
%Given two hybrid automata $\mathcal{A}_{1}$ and $\mathcal{A}_{2}$, is every timed trace in $\mathcal{A}_{1}$ also a timed trace in $\mathcal{A}_{2}$
%\end{prob}

\begin{prob}[LTL Model Checking Problem \cite{Alur2000}]
\label{prob:ltl}
Given a hybrid automaton $\mathcal{A}$ and a LTL formula $\phi$ , determine if $\mathcal{A}$ satisfies $\phi$.
\end{prob}

\begin{prob}[CTL Model Checking Problem \cite{Alur2000}]
\label{prob:ctl}
Given a hybrid automaton $\mathcal{A}$ and a CTL formula $\phi$ , determine if $\mathcal{A}$ satisfies $\phi$.
\end{prob}



\subsection{Running Example}
Figure~\ref{fig:UAV} is the hybrid automaton which will be abstracted in with each technique. Some abstractions call for a specific type of hybrid automaton, we will adjust the components accordingly and draw the new automaton.

The example hybrid system that we use is a simplified scenario of an unmanned aerial vehicle (UAV) taking off to height $H_{1}$, move forward over a distance of $M$ or ascend higher to height $H_{2}$. The UAV will not start landing until it has reached the distance $M$.
There are many different properties we can investigate in this hybrid system, we will consider the following three questions:
\begin{itemize}
    \item{Will the UAV take off, fly at height $H_{1}$ before landing?}
    \item{Will the UAV ever reach height $H_{2}$?}
    \item{Will the UAV fly forward, then go higher and then fly on?}
\end{itemize}
\begin{figure}[H]
    \begin{center}
        \begin{tikzpicture}[node distance=4cm]
            \node [state,initial] (s0) {$l_{0}$\\$\dot{x},\dot{z}=0$};
            \node [state] (s1) [above right of = s0] {$l_{1}$\\$\dot{z}=a$\\$0\leq z\leq H_{1}$};
            \node [state] (s2) [below right of = s1] {$l_{2}$\\$\dot{x}=cx$\\$0\leq x\leq M$\\$H_{1}\leq z\leq H_{2}$};
            \node [state] (s3) [below right of = s0] {$l_{4}$\\$\dot{z}=-a$\\$0\leq z\leq H_{2}$};
            \node [state] (s4) [right of = s2] {$l_{3}$\\$\dot{z}=b$\\$0\leq x \leq M$\\$H_{1}\leq z\leq H_{2}$};

            \path[->] (s0) edge node [left] {Take Off} (s1)
                (s1) edge node [right,yshift=2mm] {Move \\ $z=H_{1}$} (s2)
                (s2) edge node [right,yshift=-2mm] {Land \\ $x=M$} (s3)
                    edge [bend left] node [above] {Boost \\ $z < H_{2}$} (s4)
                (s3) edge node [left,yshift=-3mm] {Landed \\ $z=0$ \\ $x:=0$} (s0)
                (s4) edge [bend left] node [below] {Enough Height \\ $z=H_{2}$} (s2);
        \end{tikzpicture}
        \caption{Hybrid Automaton of a UAV.}
        \label{fig:UAV}
    \end{center}
\end{figure}

Formally, the hybrid automaton in Figure~\ref{fig:UAV} consists of the following:
\begin{align*}
V & = \{x,z\} \text{ with } \val(V)\subseteq\mathbb{R}^{2} \\
L & = \{l_{0},l_{1}, l_{2}, l_{3}, l_{4}\} \\
A & = \{ \text{Take Off, Move, Boost, Enough Height, Land, Landed} \} \\
Init & = \{(l_{0},(0,0))\} \\
Inv(l_{0}) & = (0,0),\ \ Inv(l_{1}) = (0,[0,H_{1}]),\ \ Inv(l_{2}) = ([0,M],[H_{1},H_{2}]),\\
Inv(l_{3}) & = ([0,M],[H_{1},H_{2}]), \ \ Inv(l_{4})=([0,M],[0,H_{2}]) \\
f(l_{0}) &=(\dot{x}=0,\dot{z}=0),\ \ f(l_{1})=(\dot{x}=0,\dot{z}=a),\ \ f(l_{2})=(\dot{x}=cx,\dot{z}=0),\\
f(l_{3}) &=(\dot{x}=0,\dot{z}=b), \ \ f(l_{4}) = (\dot{x}=0,\dot{z}=-a) \\
E & = \{e_{1}=(l_{0},\text{Take Off},l_{1}),\ e_{2}=(l_{1},\text{Move},l_{2}),\  e_{3}=(l_{2},\text{Boost},l_{3}),\\
& e_{4}=(l_{3},\text{Enough Height},l_{2}),\  e_{5}=(l_{2},\text{Land},l_{4}),\ e_{6}=(l_{4},\text{Landed},l_{0})\} \\
g(e_{1}) & =\emptyset,\ \ g(e_{2})=(\emptyset,H_{1}),\ \ g(e_{3})=\emptyset, \ \ g(e_{4})=(\emptyset,H_{2}),\ \ g(e_{5})=(M, \emptyset),\ \ g(e_{6})=(\emptyset,0) \\
r(e_{1}) & =(x,z),\ \ r(e_{2})=(x,z),\ \ r(e_{3})=(x,z),\ \ r(e_{4})=(x,z),\ \ r(e_{5}) = (x,z),\ \ r(e_{6})=(0,z) \\
\end{align*}

The three properties can be written in CTL as
\[
(x=0 \land z=0) \until (x=0 \land z>0) \until (z=H_{1}) \until (z<H_{1})
\]
\[
\exists\lozenge (z=H_{2})
\]
\[
\forall \lozenge (z>0 \until \bigcirc z=0)
\]

\subsection{Discrete Abstraction}
\label{sec:disc}
\input{discrete}


\subsection{Predicate Abstraction}
\label{sec:pred}
\input{predicate}

%\subsection{Data Abstraction}
%The larger the set of polynomials $P$ the finer the abstraction. Can we measure the probable deviation of the abstraction just knowing $P$? If so, can we impose a limit on this, to find an optimal $P$?

\subsection{Bisimulation/Simulation Abstraction}
\label{sec:bisimsim}
\input{bisimsim}


\subsection{$\mu$-Abstraction}
\label{sec:mu}
\input{mu}

%\subsection{Hierarchical Abstraction}
%Based on simulation and bisimulation.

%\subsection{Abstraction Based on Language Inclusion}

%\subsection{Compositional Abstractions}

\subsection{Abstracting Probabilistic Automata}
\label{sec:prob}
\input{probabilistic}


\subsection{Counterexample-guided abstraction refinement (CEGAR)}
\label{sec:cegar}
\input{cegar}


%\subsection{Flowpipe approximation}
%Type of approximation of the reachable set of states.

\subsection{Barrier certificates}
\label{sec:barrier}
\input{barrier}
%Used to determine upper bound of reachability of sets of states.

\subsection{Phase Portrait Approximation}
\label{sec:phase}
\input{phase}

\subsection{Other Abstractions}
\label{sec:misc}
\input{other}


\subsection{Some known hybrid automata which are undecidable}
\label{sec:undec}
\input{undecidable}


\begin{table}
\begin{center}
\begin{tabular}{|l|l|l|l|l|l|l|l|l|}\hline
Type & Discrete & Predicate & Bisimulation & $\mu$ & Probabilistic & CEGAR & Barrier & Phase Portrait \\ \hline
Rectangular &  &  &  &  &  &  &  & \\ \hline
Linear &  &  &  &  &  &  &  & \\ \hline
Timed &  &  &  &  &  &  &  & \\ \hline
Stochastic &  &  &  &  &  &  &  & \\ \hline
O-Minimal &  &  &  &  &  &  &  & \\ \hline
\end{tabular}
\label{tab:abssummary}
\caption{Summary of the presented types of hybrid systems.}
\end{center}
\end{table}
